---
title: "How to use FasStepGraph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How_to_use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(FastStepGraph)
library(MASS)     # mvrnorm

# If you directly cloned the github repository, then you should uncomment these lines to load the functions:
# source('FastStepGraph.R')
# source('SigmaAR.R')
```

Simulate Gaussian Data with an Autoregressive (AR) Model:

```{r}
set.seed(1234567)
phi = 0.4 
p = 100  # number of variables (dimension)
n = 50 # number of samples

Sigma = SigmaAR(p, phi)
Omega = solve(Sigma)  
Omega[abs(Omega) < 1e-5] = 0  

# Generate Data from a Gaussian distribution 
X = list()
X = MASS::mvrnorm(n, mu=rep(0,p), Sigma)
X = scale(X) # data normalization to zero mean and unit variance 
```

Afterwards, fit the Omega matrix by calling the Fast Step Graph function, like:

```{r}
t0 <- Sys.time() # INITIAL TIME
G = FastStepGraph(X, alpha_f = 0.22, alpha_b = 0.14)
difftime(Sys.time(), t0, units = "secs")
all.equal(Omega, G$Omega)
```

If the `nei.max` argument is omitted, it will be 5. If you don't know the `alpha_f` and `alpha_b` values, the use cross-validation. To find the optimal $\mathbf{\alpha_f}$ and $\mathbf{\alpha_b}$ parameters for the previously generated **X** data, we can perform a cross-validation on a combination grid as follows:

```{r}
t0 <- Sys.time() # INITIAL TIME
res = cv.FastStepGraph(X, data_shuffle = FALSE)
difftime(Sys.time(), t0, units = "secs")

print(res$alpha_f_opt)
print(res$alpha_b_opt)
all.equal(Omega, res$Omega)
```

The arguments `n_folds = 5`, `alpha_f_min = 0.1`, `alpha_f_max = 0.9`, `n_alpha = 32` (size of the grid search) and `nei.max = 5`, have defaults values and can be omitted. Note that, `cv.FastStepGraph(X)` is not an exhaustive grid search over $\mathbf{\alpha_f}$ and $\mathbf{\alpha_b}$. This is a heuristic that always sets $\mathbf{\alpha_b}$ = $\frac{\mathbf{\alpha_f}}{2}$. It is recommended to shuffle the rows of X before running cross-validation. The default value of `data_shuffle = TRUE`, but if you want to disable row shuffle, set it to `data_shuffle = FALSE`.

To increase time performance, you can run `cv.FastStepGraph(X)` in parallel. Next, you'll need to install and register a parallel backend. Since Windows does not support forking, the same backend that works in a Linux or OS X environment will not work in Windows. To run on a Linux system the **doParallel** dependency must be installed `install.packages("doParallel")`. To run on a Windows system, **doSNOW** is used `install.packages("doSNOW")`. These parallel packages will also require the following dependencies: **foreach**, **iterators** and **parallel**. Make sure you satisfy them. Then call the method setting the parameter **parallel = TRUE**, as follows:

```{r}
t0 <- Sys.time() # INITIAL TIME
res = cv.FastStepGraph(X, data_shuffle = FALSE, parallel = TRUE)
difftime(Sys.time(), t0, units = "secs")

print(res$alpha_f_opt)
print(res$alpha_b_opt)
all.equal(Omega, res$Omega)
```

